\mytitle

### 2021-12-14 paper.tex
adding ranges
refine ranges
refine range
remove parenthesis
shuffle terms
improve formulation

In the following an ILP-aware implementation of the real dilogarithm
is presented, which allows for the parallel execution of different
floating\-/point operations.  The real dilogarithm is defined as
%
\begin{align}
  \Li(x) &= -\int_0^x \frac{\ln(1-t)}{t}dt, &
  &(x<1).
\end{align}
%
For $|x|<1$ the real dilogarithm has the series representation
%
\begin{align}
  \Li(x) = \sum_{k=1}^\infty \frac{x^k}{k^2}.
  \label{eq:series}
\end{align}
%
Further series representations exist, for example the accelerated
series
%
\begin{align}
  \Li(x) = \sum_{k=0}^\infty \frac{B_k}{(k+1)!} [-\ln(1-x)]^{k+1},
\end{align}
%
where $B_k$ are the Bernoulli numbers.  A series representation in
terms of Chebyshev polynomials of the first kind $T_n(x)$ is given in
\cite{luke},
%
\begin{align}
  \Li(x) = \sum_{k=0}^\infty a_n T_n(2x-1),
  \label{eq:cheby}
\end{align}
%
which is used for example in the ROOT program library \cite{root}.
The r.h.s.\ of Eqs.~\eqref{eq:series}--\eqref{eq:cheby} can be summed,
for example, using Horner's scheme or Clenshaw's algorithm
\cite{clenshaw}.  These summation techniques are, however, purely
sequential and thus do not make use of ILP.  In the following an
approximation of the real dilogarithm is presented, which makes use of
ILP, using a rational function approximation in combination with
Estrin's scheme, similar to the approach used in
Eq.~\eqref{eq:mixed}.%
\footnote{The Cephes library \cite{cephes} implements an equivalent
  rational function approximation for Spence's function
  $\operatorname{Sp}(x)=\Li(1-x)$.}

We define the real dilogarithm $\Li(x)$ for all $x\in\mathbb{R}$ as
the real part of its complex continuation.  The real dilogarithm can
be mapped onto the interval $[0,1/2]$ by using the following
identities \cite{lewin}:
%
\begin{align}%
  \Li(x) ={}& \Li\left(\frac{1}{1-x}\right) - \frac{\pi^2}{6} \notag \\ &+ \ln(1-x)\left[\frac{1}{2}\ln(1-x) - \ln(-x)\right], \\ % (x<-1), \\
  \Li(x) ={}& -\Li\left(\frac{x}{x-1}\right) - \frac{1}{2}\ln^2(1-x), \\ % (-1<x<0), \\
  \Li(x) ={}& -\Li(1-x) + \frac{\pi^2}{6} - \ln(x) \ln(1-x), \\ % (\frac{1}{2}\leq x<1), \\
  \Li(x) ={}& \Li\left(1-\frac{1}{x}\right) + \frac{\pi^2}{6} \notag \\ &- \ln(x)\left[\ln\left(1-\frac{1}{x}\right) + \frac{1}{2}\ln(x)\right], \\ % (1<x<2), \\
  \Li(x) ={}& -\Li\left(\frac{1}{x}\right) + \frac{\pi^2}{3} - \frac{1}{2} \ln^2(x), % (x\geq 2),%
\end{align}%
%
and by using the special values $\Li(-1)=-\pi^2/12$, $\Li(0)=0$,
$\Li(1/2)=\pi^2/12-\ln^2(2)/2$, $\Li(1)=\pi^2/6$ and $\Li(2)=\pi^2/4$.
Within the interval $[0,1/2]$ the real dilogarithm has the series
expansion \eqref{eq:series}, which can be written as
%
\begin{align}
  \Li(x) = x \sum_{k=0}^\infty \frac{x^{k}}{(k+1)^2}.
  \label{eq:series2}
\end{align}
%
We approximate the sum on the r.h.s.\ of Eq.~\eqref{eq:series2} using
a rational minimax approximant on the interval $[0,1/2]$ using the
\texttt{MiniMaxApproximation} from Wolfram/Mathematica
\cite{mathematica},
%
\begin{align}
  \sum_{k=0}^\infty \frac{x^{k}}{(k+1)^2} \approx
  \frac{\sum_{k=0}^5 p_kx^k}{\sum_{k=0}^6 q_kx^k}.
  \label{eq:minimax}
\end{align}
%
The coefficients $p_k$ and $q_k$ are listed in \tabref{tab:coeffs}.
%
\begin{table}[t]
  \centering
  \caption{Coefficients of the numerator and denominator polynomials
    for the minimax approximant \eqref{eq:minimax}.}
  \begin{tabular}{lr}
    \toprule
    $p_0$ & $ 0.9999999999999999502\cdot 10^{+0}$ \\
    $p_1$ & $-2.6883926818565423430\cdot 10^{+0}$ \\
    $p_2$ & $ 2.6477222699473109692\cdot 10^{+0}$ \\
    $p_3$ & $-1.1538559607887416355\cdot 10^{+0}$ \\
    $p_4$ & $ 2.0886077795020607837\cdot 10^{-1}$ \\
    $p_5$ & $-1.0859777134152463084\cdot 10^{-2}$ \\
    $q_0$ & $ 1.0000000000000000000\cdot 10^{+0}$ \\
    $q_1$ & $-2.9383926818565635485\cdot 10^{+0}$ \\
    $q_2$ & $ 3.2712093293018635389\cdot 10^{+0}$ \\
    $q_3$ & $-1.7076702173954289421\cdot 10^{+0}$ \\
    $q_4$ & $ 4.1596017228400603836\cdot 10^{-1}$ \\
    $q_5$ & $-3.9801343754084482956\cdot 10^{-2}$ \\
    $q_6$ & $ 8.2743668974466659035\cdot 10^{-4}$ \\
    \bottomrule
  \end{tabular}
  \label{tab:coeffs}
\end{table}

>>>>

In the following an ILP-aware implementation of the real dilogarithm
is presented, which allows for the parallel execution of different
floating\-/point operations.  The real dilogarithm is defined as
%
\begin{align}
  \Li(x) &= -\int_0^x \frac{\ln(1-t)}{t}dt, &
  &x<1.
\end{align}
%
For $|x|<1$ the real dilogarithm has the series representation
%
\begin{align}
  \Li(x) = \sum_{k=1}^\infty \frac{x^k}{k^2}.
  \label{eq:series}
\end{align}
%
Further series representations exist, for example the accelerated
series
%
\begin{align}
  \Li(x) = \sum_{k=0}^\infty \frac{B_k}{(k+1)!} [-\ln(1-x)]^{k+1},
\end{align}
%
where $B_k$ are the Bernoulli numbers.  A series representation in
terms of Chebyshev polynomials of the first kind $T_n(x)$ is given in
\cite{luke},
%
\begin{align}
  \Li(x) = \sum_{k=0}^\infty a_n T_n(2x-1),
  \label{eq:cheby}
\end{align}
%
which is used for example in the ROOT program library \cite{root}.
The r.h.s.\ of Eqs.~\eqref{eq:series}--\eqref{eq:cheby} can be summed,
for example, using Horner's scheme or Clenshaw's algorithm
\cite{clenshaw}.  These summation techniques are, however, purely
sequential and thus do not make use of ILP.  In the following an
approximation of the real dilogarithm is presented, which makes use of
ILP, using a rational function approximation in combination with
Estrin's scheme, similar to the approach used in
Eq.~\eqref{eq:mixed}.%
\footnote{The Cephes library \cite{cephes} implements an equivalent
  rational function approximation for Spence's function
  $\operatorname{Sp}(x)=\Li(1-x)$.}

We define the real dilogarithm $\Li(x)$ for all $x\in\mathbb{R}$ as
the real part of its complex continuation.  The real dilogarithm can
be mapped onto the interval $[0,1/2]$ by using the following
identities \cite{lewin}:
%
\begin{align}
  \Li(x) ={}& \ln(1-x)\left[\frac{1}{2}\ln(1-x) - \ln(-x)\right] - \frac{\pi^2}{6}
              \notag \\ &+\Li\left(\frac{1}{1-x}\right), \quad x\leq -1, \\
  \Li(x) ={}& -\Li\left(\frac{x}{x-1}\right) - \frac{1}{2}\ln^2(1-x),
              \notag \\ & -1<x<0, \\
  \Li(x) ={}& -\Li(1-x) + \frac{\pi^2}{6} - \ln(x) \ln(1-x),
              \notag \\ &\frac{1}{2}<x<1, \\
  \Li(x) ={}& \frac{\pi^2}{6} - \ln(x)\left[\ln\left(1-\frac{1}{x}\right) + \frac{1}{2}\ln(x)\right]
              \notag \\ &+ \Li\left(1-\frac{1}{x}\right), \quad 1<x\leq 2, \\
  \Li(x) ={}& -\Li\left(\frac{1}{x}\right) + \frac{\pi^2}{3} - \frac{1}{2} \ln^2(x), \quad x>2,
\end{align}
%
and by using the special values $\Li(-1)=-\pi^2/12$, $\Li(0)=0$,
$\Li(1/2)=\pi^2/12-\ln^2(2)/2$, $\Li(1)=\pi^2/6$ and $\Li(2)=\pi^2/4$.
Within the interval $[0,1/2]$ the real dilogarithm has the series
expansion \eqref{eq:series}, which can be written as
%
\begin{align}
  \Li(x) = x \sum_{k=0}^\infty \frac{x^{k}}{(k+1)^2}.
  \label{eq:series2}
\end{align}
%
We approximate the sum on the r.h.s.\ of Eq.~\eqref{eq:series2} in the
form of a rational minimax approximant on the interval $[0,1/2]$ using
the \texttt{MiniMaxApproximation} from Wolfram/Mathematica
\cite{mathematica},
%
\begin{align}
  \sum_{k=0}^\infty \frac{x^{k}}{(k+1)^2} \approx
  \frac{\sum_{k=0}^5 p_kx^k}{\sum_{k=0}^6 q_kx^k}.
  \label{eq:minimax}
\end{align}
%
The coefficients $p_k$ and $q_k$ are listed in \tabref{tab:coeffs}.
%
\begin{table}[t]
  \centering
  \caption{Coefficients of the numerator and denominator polynomials
    for the minimax approximant \eqref{eq:minimax}.}
  \begin{tabular}{lr}
    \toprule
    $p_0$ & $ 0.9999999999999999502\cdot 10^{+0}$ \\
    $p_1$ & $-2.6883926818565423430\cdot 10^{+0}$ \\
    $p_2$ & $ 2.6477222699473109692\cdot 10^{+0}$ \\
    $p_3$ & $-1.1538559607887416355\cdot 10^{+0}$ \\
    $p_4$ & $ 2.0886077795020607837\cdot 10^{-1}$ \\
    $p_5$ & $-1.0859777134152463084\cdot 10^{-2}$ \\
    $q_0$ & $ 1.0000000000000000000\cdot 10^{+0}$ \\
    $q_1$ & $-2.9383926818565635485\cdot 10^{+0}$ \\
    $q_2$ & $ 3.2712093293018635389\cdot 10^{+0}$ \\
    $q_3$ & $-1.7076702173954289421\cdot 10^{+0}$ \\
    $q_4$ & $ 4.1596017228400603836\cdot 10^{-1}$ \\
    $q_5$ & $-3.9801343754084482956\cdot 10^{-2}$ \\
    $q_6$ & $ 8.2743668974466659035\cdot 10^{-4}$ \\
    \bottomrule
  \end{tabular}
  \label{tab:coeffs}
\end{table}
